---
title: "CLT"
editor_options: 
  chunk_output_type: console
---

## CLT in action

$X_{1}, X_{2},\ldots$ are IID $\text{Bern}(1/4)$ random variables, where $X_{i}$ is indicator of Heads. Then $\bar{X}_{n}$ is the proportion of Heads in our $n$ tosses.

We will see the CLT approximation to $\bar{X}_{n}$ and $\sum_{i=1}^{n} X_{i}$ as $n$ increases from 5 through 300. Remember, CLT tells us that

$$
\bar{X}_{n} \overset{\cdot}{\sim} N\left(0.25, \frac{0.25(0.75)}{n}\right) \qquad \text{ and } \qquad  \sum_{i=1}^{n} X_{i} \overset{\cdot}{\sim} N\left(0.25 n, 0.25(0.75)n \right)
$$

```{r echo = F, message = F, fig.align="center"}
library(tidyverse)
set.seed(20)
B <- 300
J <- 500
p <- 0.25
n_vec <- c(5,10, 20,50, 100, 200, B)
N <- length(n_vec)
props <- matrix(NA, nrow = N, ncol = J )
for(j in 1:J){
  x <- rbinom(B, 1, p)
  for(i in 1:N){
    props[i,j] <- mean(x[1:n_vec[i]])
  }
}

plot_ls <- list()
for(n in 1:N){
  plot_ls[[n]] <- data.frame(xbar = props[n,], 
             n = n_vec[n],
             sum_x = props[n,]*n_vec[n])
}

vls <- data.frame(vals = c(n_vec * p, rep(p, N)),
           RV = c(rep("sum_x", N), rep("xbar", N)),
           n = factor(rep(n_vec, 2)))
do.call(rbind, plot_ls) |>
  mutate(n = factor(n)) |>
  pivot_longer(cols = c(xbar, sum_x), names_to = "RV",
               values_to = "vals") |>
  ggplot(aes(x = vals)) +
  geom_histogram(bins = 40) +
  geom_vline(data = vls, aes(xintercept = vals), col = "orange") +
    facet_grid(n ~ RV, scales = "free")  +
  theme_minimal() +
  labs(x = "Values")+
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        text=element_text(size = 14))


```

## CLT continuity correction

```{r echo = F}
# Parameters
n <- 20
p <- 0.25

x <- 0:n
binom_df <- data.frame(
  x = x,
  prob = dbinom(x, size = n, prob = p)
)



# Continuous normal approximation
mu <- n * p
sigma <- sqrt(n * p * (1 - p))

normal_df <- data.frame(
  x = seq(0, n, length.out = 1000)
)

normal_df$density <- dnorm(normal_df$x, mean = mu, sd = sigma)

# Continuity-corrected region for P(a <= X <= b)
a <- 2
b <- 7

```

Suppose we have $Y \sim \text{Binom}(`r n`, `r p`)$ and we'd like to approximate $P(`r a` \leq Y \leq `r b`)$ using the CLT. Visually:


```{r echo = F}
shade_df <- data.frame(
  x = seq(a - 0.5, b + 0.5, length.out = 500)
)

shade_df$y <- dnorm(shade_df$x, mean = mu, sd = sigma)

# Plot
ggplot() +
  
  # Binomial bars
  geom_col(
    data = binom_df,
    aes(x = x, y = prob),
    fill = "blue",
    alpha = 0.5,
    color = "white",
    width = 0.9
  ) +
  
  # Normal approximation curve
  geom_line(
    data = normal_df,
    aes(x = x, y = density),
    color = "purple",
    linewidth = 1.2
  ) +
  
  # Continuity correction shaded area
  geom_area(
    data = shade_df,
    aes(x = x, y = y),
    fill = "magenta",
    alpha = 0.4
  ) +
  
  # Vertical line at a
  geom_vline(
    xintercept = a,
    linetype = "dashed",
    color = "blue",
    linewidth = 1
  ) +

  # Vertical line at b
  geom_vline(
    xintercept = b,
    linetype = "dashed",
    color = "blue",
    linewidth = 1
  ) +
  
  # Vertical line at b + 0.5 (continuity correction)
  geom_vline(
    xintercept = b + 0.5,
    linetype = "dashed",
    color = "magenta",
    linewidth = 1
  ) +
  # Vertical line at a - 0.5 (continuity correction)
  geom_vline(
    xintercept = a -  0.5,
    linetype = "dashed",
    color = "magenta",
    linewidth = 1
  ) +
  
  
  annotate(
    "text",
    x = b + 3 ,
    y = 0.17,
    label = paste("Approximation without correction"),
    color = "blue",
    hjust = -0.1
  ) +
  
  annotate(
    "text",
    x = b + 3,
    y = 0.15,
    label = "Approximation with correction",
    color = "magenta",
    hjust = -0.1
  ) +
  
  labs(
    title = "CLT: Continuity Correction",
    x = "y",
    y = "Probability / Density"
  ) +
  
  theme_minimal(base_size = 14)
```

Let's look at the probabilities themselves:


```{r}
# exact using Binomial CDF
pbinom(b, n, p) - pbinom(a-1, n , p)

# CLT without correction
mu_y <- n*p
sd_y <- sqrt(n *p * (1-p))
pnorm((b - mu_y)/ sd_y) - pnorm((a - mu_y)/sd_y) 

# CLT with correction
pnorm((b + 0.5 - mu_y)/ sd_y) - pnorm((a -0.5 - mu_y)/sd_y) 
```


