---
title: "Intro to R and Simulations"
author: "Your name"
date: "02-18-2026"
output:
  html_document: default
editor_options: 
  chunk_output_type: console
---

# R is a calculator

```{r}

```

R respects PEMDAS:

```{r}

```

# R objects

### Booleans

Booleans/logicals can be formed with (in)equality symbols:

```{r}

```

### Variables

Storing/saving values into variables is achieved by using the syntax `<variable name> <- <value>`. Note that the variable name cannot begin with a number! The result of a stored variable is not displayed to the user. Type out the variable name after storing to explicitly show the values.

```{r}

```

### Vectors

Create vectors of values (i.e. a list of values) using the concatenate `c()` function:

```{r}

```

The syntax `a:b` creates a sequence of integers from `a` to `b`.

```{r}

```

R is vectorized language, so it operates element-wise in vectors. Also, under the hood, booleans/logicals are treated as 1/0 for TRUE/FALSE

```{r}

```

# Functions

A lot of the commands we have been working with are **functions**. You can tell a command is a function if it has parentheses (e.g. `sum()` and `c()`). A function usually takes in some argument(s) as input and returns something back as output. We can write custom functions using the function `function()`. We need to specify the name of the function, what it expects as input, as well as what to return back to the user.

### Birthday problem, generalized

Suppose a group of $k$ people each choose a number randomly with replacement from a list of $n$ distinct numbers. If $k \leq n$, the probability of at least one match in the group is 1 minus the probability of no matches:

$$1 - \frac{n(n-1)(n-2)\cdots(n-k+1)}{n^k} = 1- \frac{n!}{(n-k)!n^k} =  1- \frac{n!k!}{k!(n-k)!n^k} = 1 -\binom{n}{k} \frac{k!}{n^k}$$ We can write this a function in R to obtain this probability for arbitrary $n$ and $k$:

```{r}

```

Note that the function name is prob_match, and the arguments/inputs are called `n` and `k`. The output is the probability of at least one match for the specified values of `n` and `k`. We can now use this function! For example, the probability of at least one match among 23 people choosing a number between 1 and 365 is:

```{r}

```

# Basic plotting

We can plot one variable against another (e.g. input and output of a function) using the `plot()` function. The first argument is a vector of values for the x-axis variable (e.g. input), and the second argument is a vector for the y-axis variable (e.g. output).

For example, we will plot probabilities of each least one birthday match in a room full of $k = 1,\ldots, 50$ people:

```{r fig.height=4}

```

# Sampling and simulating

The `sample()` function generates random values from a pool of values contained in a specific vector. You can twist many knobs to get sampling with or without replacement, uneven probabilities, and different sized samples.

```{r}

```

# Approximating probabilities

In the following, we simulate flipping a fair coin once:

```{r}

```

We can approximate probabilities by repeatedly simulating an experiment, counting up how many times a favorable outcome occurred, and dividing by the total number of experiments we did. This is easily done using the `replicate()` function:

```{r}

```
